Title here

### 2016-12-06 portfolio/text/ms.Rmd
Rewrap and edit paper

A fundamental challenge for individuals who rely on the extraction of natural
resources for their livelihood is income variability.
Fishers, loggers, farmers, and miners, among others,
face income uncertainty driven by a mix of
environmental, biological, and economic variability.
A key component underlying the sustainability
of these industries is identifying both
how regulatory actions and individual behaviour affect income variability
and by extension the economic stability and
sustainable use of natural resources.
For commercial fisheries, recent work has identified that
targeting a diversity of species can reduce income variability.
However, any diversity-variability relationship
is embedded in a regulatory framework that restricts individual ability to
change fishing behaviour. Here we identify three primary axes by which fishers
in Alaska, one of the world's most productive fishery regions, can potentially
diversify fishing activities and investigate how individual actions along each
diversity axis affect both revenue and revenue variability. We construct a novel
statistical model to simultaneously assess the value to individual fishers of
choosing to (1) own permits that allow targeting a diverse fish species
assemblage, (2) diversify by owning multiple permits, or (3) catch a greater
species diversity within the constraints of a given permit.
We show contrasting effects of diversity on
revenue and revenue variability along the three axes.
First, we show that fishing permits with
higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity).
Next, we show that fishing an additional permit
is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits).
Finally, we show that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue the following year.
Thus, while there are clear benefits
to some axes of diversification, others are less effective.
Importantly, the most effective measure for
increasing revenue and decreasing risk ---
purchasing additional permits to participate in additional fisheries ---
was also the most costly.
At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies,
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

@@

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in geographical areas to buffer against disease outbreak, wildfires, and
price variation REF. However, otherwise well-intentioned regulations can limit
the actions available to individuals to reduce income variability, or
incentivize against diversification. For example, in terrestrial agriculture,
crop subsidies in the US may incentivize some farms to specialize [e.g. dairy,
grains, soybeans; @mishra2002]. In other instances, external market forces
(prices, demand) may discourage the adoption of diverse strategies and help
promote specialization [@bradshaw2004].

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture (@sethi2012b). The majority of insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
available to commercial fishers [@kasperski2013] and both fish and fishers are
spatially mobile. However, fishers still have the ability to lower risk. They
can, for example, target multiple species as permitted by a license, own
multiple licenses, or diversify spatially and fish in multiple regions. The
effectiveness of these diversification strategies for improving revenue and
reducing revenue variability remains uncertain. Furthermore, theory and
empirical studies both suggest there can be substantial benefits to
specialization (REFS). The majority of papers that examine effects of
diversification on revenue do so at aggregate levels (e.g. households, counties,
communities; REFS), rather than focus on individuals. Because of this, key
uncertainties remain about the fundamental relationship between diversity,
fishing revenue, and fishing revenue variability at the level of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability for those whose livelihood depends on the extraction of natural
resources. Fisheries in Alaska are among the highest-valued fisheries in the
world with 2.7 billion kg landed biomass valued at nearly 2 billion USD annually
[@nmfs2016]. The industry supports a range of fishing fleets ranging from small,
open, single-operator vessels to multi-million-dollar state-of-the-art ships
with large crews. Like much of the rest of the world [@hilborn2003a], the days
of open, unregulated fisheries in most developed countries have passed ---
fishers must own a license to participate in virtually all fisheries. And while
such regulations have proven helpful in reducing or eliminating overfishing,
they limit the available strategies to reduce income variability. For example,
most permits limit fishing to a single species, diversifying by purchasing
additional fishing permits can be expensive (e.g. ~\$XX for king crab), and some
fisheries limit the entry of new participants through individual fishing quotas
(IFQs) (REF).

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals rather than
vessels we can track people that switch fishery sectors over time, or alter
strategies and use multiple vessels. We develop and apply a novel statistical
approach for understanding the effect of diversity on both revenue and revenue
variability that builds on results from vessel [@kasperski2013], fleet
[@sethi2012b], and community-level variability [@sethi2014]. Our model controls
for fishing effort and latent time effects (e.g. changing environmental
conditions), simultaneously estimates the effect of species diversity on
individual revenue and revenue variability, and estimates effects both within
and across fishing strategies (sets of similar licenses or permits). To our
knowledge, this is the first application of this type of model in any natural
resource field.

@@

Overview paragraph here.

@@

We obtained fisheries landings and revenue data for
all permit holders in Alaska
from `r range_yrs[1]` to `r range_yrs[2]`
from the Alaskan Commercial Fisheries Entry Commission (CFEC).
These data represent a total of
nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD
and a combined `r round(ndat$initial/1e6,1)` million rows of data.
We adjusted for inflation by setting all revenues in 2009 USD
based on data from TODO.
We implemented a number of filtering steps
to focus on individuals actively engaged in commercial fishing
(Supporting Information).
In particular, we removed permit holders whose median revenue
was less than \$10,000 2009 USD.

@@

Our models focused on species diversity
as a possible predictor of revenue and revenue variability
while controlling for effort.
For consistency with previous analyses [e.g. @kasperski2013],
we calculated effective species diversity for
permit holder $i$ and time (year) $t$ as the inverse of
Simpson's diversity index $\lambda$
(\citeauthor{simpson1949} \citeyear{simpson1949},
\citeauthor{jost2006} \citeyear{jost2006};
also referred to in economics as the Herfindahl-Hirschman Index)
weighted by revenue $R$:

@@

To jointly model revenue and revenue variability as a function of
individual-level covariates, we use a variance function regression model, which
separately models change in the mean and variance as
a function of covariates [e.g. @davidian1987; @hall1989]. 
Whereas traditional linear models or 
hierarchical linear models generally assume
homoscedasticity --- that the residual variance 
is the same for all values of the predictor variables ---
the variance-function regression model allows the variance to be heteroscedastic. 
For application to economic data on individual fishing
portfolios, this type of model allows certain strategies and levels of
diversification to be more or less variable than others.

We extended the basic variance-function regression to a hierarchical
Bayesian variance-function regression model with Gaussian errors
predicting $\log$ revenue, $\log(R_{i,t})$, for
each fisher $i$ and time $t$:

@@

\noindent
Each year $t$ of revenue for fisher $i$
is assigned a strategy indexed by $j$
(Supporting Information, Table \ref{tab:strategies})
and expected $\log(R_{i,t})$ is modelled as:
<!-- Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid -->
<!-- potential confusion like using the term "expected log(R)".   -->
<!-- I also just don't like hats... but that's my problem. -->

@@

\bibliographystyle{apalike}
\bibliography{refs}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}
\caption{
Trends in revenue and diversification over time.
(A) Map of fishing regions in Alaska, USA.
(B) Total revenue in billions of 2009 U.S. dollars from `r range_yrs[1]` to
`r range_yrs[2]` for multiple (grey) and single (blue) permit holders.
(C) Proportion of single permit holders for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the
`r n_jacknifed` top species groups (thin lines).
(D) Mean effective diversity of species fished (Simpson's diversity index)
for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the
`r n_jacknifed` top species groups (thin lines).
}
\label{fig:ts}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}
\caption{Individuals who fish
permits and permit combinations (permit strategies) with higher species diversity
have lower expected revenue variability.
On the y-axis, dots, thick lines, and thin lines represent posterior medians,
50\%, and 95\% credible intervals for
the `r nstr` most common permit strategies.
Thick black line, and dark and light grey shaded regions
indicate median, 50\%, and 95\% credible intervals
of strategy-level regression built into the hierarchical model.
Estimated variability represents expected standard deviation
for a permit holder who does not change
species diversity or days fished from year to year.
}
\label{fig:across}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
\caption{Diversifying permits reduces risk and usually increases revenue.
Two examples of expected revenue variability
and median revenue for
owners of (A) individual halibut and (B) gillnet herring roe permits
(dark blue)
compared with owners of the same permit plus one more (medium blue)
or two more permits (light blue).
On the y-axis, dots and lines represent posterior median and 95\%
credible intervals of expected standard deviation.}
\label{fig:spokes}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
\caption{Effects of (A) generalizing and (B) specializing from year to year for
individual permit holders across the `r nstr` most common strategies.
Changing species diversity from year to year
is associated with higher variability
for individuals within the strategies above the zero line.
Generalizing (A) or specializing (B) is associated with
greater expected revenue to the right of the zero line
and lower expected revenue to the left of the zero line.
These effects are estimated holding
fishing season length constant from year to year.
Dots and line segments represent medians and
50\% credible intervals of posteriors.
}
\label{fig:within}
\end{center}
\end{figure}

@@

>>>>

A fundamental challenge for individuals who rely on the extraction of natural
resources for their livelihood is income variability. Fishers, loggers, farmers,
and miners, among others, face high levels of income uncertainty driven by a mix
of environmental, biological, and economic variability. A key component
underlying the sustainability of these industries is identifying both how
regulatory actions and individual behaviour affect income variability and by
extension the economic stability and sustainable use of natural resources. For
commercial fisheries, recent work has identified that targeting a diversity of
species can reduce income variability. However, any diversity-variability
relationship is embedded in a regulatory framework that restricts individual
ability to change fishing behaviour. Here we identify three primary axes by
which fishers in Alaska, one of the world's most productive fishery regions, can
potentially diversify fishing activities and investigate how individual actions
along each diversity axis affect both revenue and revenue variability. We
construct a novel statistical model to simultaneously assess the value to
individual fishers of choosing to (1) own permits that allow targeting a diverse
fish species assemblage, (2) diversify by owning multiple permits, or (3) catch
a greater species diversity within the constraints of a given permit. We show
contrasting effects of diversity on revenue and revenue variability along the
three axes. First, we show that fishing permits with higher species diversity
reduces variability (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction
per unit increase in mean effective species diversity). Next, we show that
fishing an additional permit is associated with higher revenue and lower
variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
IQR] lower expected standard deviation for halibut plus another permit vs.
single halibut permits). Finally, we show that increasing species diversity
within the constraints of a permit set has a fishery-dependent effect on revenue
and is usually (`r g2_dens`% probability) associated with increased uncertainty
in revenue the following year. Thus, while there are clear benefits to some axes
of diversification, others are less effective. Importantly, the most effective
measure for increasing revenue and decreasing risk --- purchasing additional
permits to participate in additional fisheries --- was also the most costly. At
the same time, Alaskan commercial fisheries have been increasingly focussed on
single-permit (often salmon) fishing strategies. Our results suggest that the
benefit of species diversity for revenue variability is mostly an across-permit
effect, not a benefit that can be easily harnessed by an individual permit
holder. Effective methods of diversification --- purchasing additional permits
or switching permits --- are costly and often limited by regulations such as
limited entry programs and individual transferable quotas (ITQs). Take home ...

@@

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in geographical areas to buffer against disease outbreak, wildfires, and
price variation REF. However, otherwise well-intentioned regulations can limit
the actions available to individuals to reduce income variability, or
incentivize against diversification. For example, in terrestrial agriculture,
crop subsidies in the US may incentivize some farms to specialize (e.g. dairy,
grains, soybeans; @mishra2002). In other instances, external market forces
(prices, demand) may discourage the adoption of diverse strategies and help
promote specialization [@bradshaw2004].

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
available to commercial fishers [@kasperski2013] and both fish and fishers are
spatially mobile. However, fishers still have the ability to lower risk. They
can, for example, target multiple species as permitted by a license, own
multiple licenses, or diversify spatially and fish in multiple regions. The
effectiveness of these diversification strategies for improving revenue and
reducing revenue variability remains uncertain. Furthermore, theory and
empirical studies both suggest there can be substantial benefits to
specialization (REFS). The majority of papers that examine effects of
diversification on revenue do so at aggregate levels (e.g. households, counties,
communities; REFS), rather than focus on individuals. Because of this, key
uncertainties remain about the fundamental relationship between diversity,
fishing revenue, and fishing revenue variability at the level of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability for those whose livelihood depends on the extraction of natural
resources. Fisheries in Alaska are among the highest-valued fisheries in the
world with 2.7 billion kg landed biomass valued at nearly 2 billion USD annually
[@nmfs2016]. The industry supports a range of fishing fleets ranging from small,
open, single-operator vessels to multi-million-dollar state-of-the-art ships
with large crews. Like much of the rest of the world [@hilborn2003a], the days
of open, unregulated fisheries in most developed countries have passed ---
fishers must own a license to participate in virtually all fisheries. And while
such regulations have proven helpful in reducing or eliminating overfishing,
they limit the available strategies to reduce income variability. For example,
many permits limit fishing to a single species, diversifying by purchasing
additional fishing permits can be expensive (e.g. ~\$XX for king crab), and some
fisheries limit the entry of new participants through individual fishing quotas
(IFQs) (REF).

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals rather than
vessels we can track people that switch fishery sectors over time, or alter
strategies and use multiple vessels. We develop and apply a hierarchical
Bayesian model to understand the effect of diversity on both individual revenue
and revenue variability that builds on results from vessel [@kasperski2013],
fleet [@sethi2012b], and community-level variability [@sethi2014]. Our model
controls for fishing effort and latent time effects (e.g. changing environmental
conditions), simultaneously estimates the effect of species diversity on
individual revenue and revenue variability, and estimates effects both within
and across fishing strategies (sets of similar licenses or permits). To our
knowledge, this is the first application of this type of model in any natural
resource field.

@@

We developed a hierarchical variance-function regression model, which separately
models changes in the mean and variance of individual revenue as a function of
covariates [@davidian1987; @hall1989]. Whereas traditional hierarchical linear
models assume homoscedasticity --- that the residual variance does not vary
systematically --- the variance-function regression model allows the variance to
be heteroscedastic and vary as a function of covariates. For application to
economic data on individual fishing portfolios, this model allows certain
strategies and levels of diversification to be more or less variable than
others. Our model includes the number of days fished to represent fishing
effort,

@@

We obtained fisheries landings and revenue data for all permit holders in Alaska
from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan Commercial Fisheries
Entry Commission (CFEC). These data represent a total of nearly `r 
round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r 
round(ndat$initial/1e6,1)` million rows of data. We adjusted for inflation by
setting all revenues in 2009 USD [@bea2016]. We implemented a number of
filtering steps to focus on individuals actively engaged in commercial fishing
(Supporting Information). In particular, we removed permit holders whose median
revenue was less than \$10,000 2009 USD.

@@

Our models focused on species diversity as a possible predictor of revenue and
revenue variability while controlling for effort. For consistency with previous
analyses [@kasperski2013], we calculated effective species diversity for
permit holder $i$ and time (year) $t$ as the inverse of Simpson's diversity
index $\lambda$ [@simpson1949; @jost2006] (also referred to in economics as the
Herfindahl-Hirschman Index) weighted by revenue $R$:

@@

To jointly model revenue and revenue variability as a function of
individual-level covariates, we extended the basic variance-function regression
to a hierarchical Bayesian variance-function regression model with Gaussian
errors predicting $\log$ revenue, $\log(R_{i,t})$, for each fisher $i$ and time
$t$:

@@

\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy indexed
by $j$ (Supporting Information, Table \ref{tab:strategies}) and expected
$\log(R_{i,t})$ is modelled as:

<!-- Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid -->
<!-- potential confusion like using the term "expected log(R)".   -->
<!-- I also just don't like hats... but that's my problem. -->

@@

\bibliographystyle{pnas2011}
\bibliography{refs}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}

\caption{ Trends in revenue and diversification over time. (A) Map of fishing
regions in Alaska, USA. (B) Total revenue in billions of 2009 U.S. dollars from
`r range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue)
permit holders. (C) Proportion of single permit holders for all fisheries (thick
black line) and with permit holders omitted each year who fished any of the `r 
n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
species fished (Simpson's diversity index) for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the `r n_jacknifed`
top species groups (thin lines). } \label{fig:ts} \end{center} \end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}

\caption{Individuals who fish permits and permit combinations (permit
strategies) with higher species diversity have lower expected revenue
variability. On the y-axis, dots, thick lines, and thin lines represent
posterior medians, 50\%, and 95\% credible intervals for the `r nstr` most
common permit strategies. Thick black line, and dark and light grey shaded
regions indicate median, 50\%, and 95\% credible intervals of strategy-level
regression built into the hierarchical model. Estimated variability represents
expected standard deviation for a permit holder who does not change species
diversity or days fished from year to year. }

\label{fig:across}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}

\caption{Diversifying permits reduces risk and usually increases revenue. Two
examples of expected revenue variability and median revenue for owners of (A)
individual halibut and (B) gillnet herring roe permits (dark blue) compared with
owners of the same permit plus one more (medium blue) or two more permits (light
blue). On the y-axis, dots and lines represent posterior median and 95\%
credible intervals of expected standard deviation.}

\label{fig:spokes}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}

\caption{Effects of (A) generalizing and (B) specializing from year to year for
individual permit holders across the `r nstr` most common strategies. Changing
species diversity from year to year is associated with higher variability for
individuals within the strategies above the zero line. Generalizing (A) or
specializing (B) is associated with greater expected revenue to the right of the
zero line and lower expected revenue to the left of the zero line. These effects
are estimated holding fishing season length constant from year to year. Dots and
line segments represent medians and 50\% credible intervals of posteriors. }

\label{fig:within}
\end{center}
\end{figure}

@@

<!-- vim: textwidth=80 
-->
